---
output:
  github_document:
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Karlsson2013inR

<!-- badges: start -->
<!-- badges: end -->

This package implements the algorithms in Karlsson (2013) in R.

Karlsson, S. (2013). Forecasting with Bayesian Vector Autoregression. In: Elliott, G. and Timmerman, A. (eds) *Handbook of Economic Forecasting*. Elsevier B.V. Vol 2, Part B., pp. 791-897.

## Installation

```{r, eval=FALSE}
remotes::install_github("markjwbecker/Karlsson2013inR", force = TRUE, upgrade = "never")
```

## Introduction

The BVAR model is


$$
\begin{aligned}
y_t'&=\sum_{i=1}^p y_{t-i}' A_i + x_t' C + u_t'\\
&=z_t'\Gamma + u_t'
\end{aligned}
$$

where $x_t$ is a vector of $d$ deterministic variables

$$
z_t' = \begin{pmatrix}y_{t-1}',\dots,y_{t-p}',x_t'\end{pmatrix}
$$

is a $k=mp + d$ dimensional vector and


$$
\Gamma= \begin{pmatrix}A_1',\dots,A_p',C'\end{pmatrix}'
$$


is a $k \times m$ matrix.
We have normally distributed errors $u_t \sim N(0, \Psi)$.

## Algorithm 4 (steady-state BVAR)

Let $A(L)= I-A_1'L-\ldots-A_p'L^p$ we can then write a stationary VAR as


$$
A(L)y_t = C'x_t +u_t
$$

The unconditional expectation is the $E(y_t)=\mu_t=A^{-1}(L)C'x_t=\Lambda x_t$. We can further rewrite the model in mean deviation form


$$
A(L)(y_t-\Lambda x_t) = u_t
$$

We can further rewrite this as a non-linear regression


$$
y_t' =x_t'\Lambda' + \left[w_t'-q_t'(I_p \otimes \Lambda') \right]\Gamma_d +u_t'
$$

where $w_t'=(y_{t-1}',\dots,y_{t-p}')$ is a $kp$-dimensional vector of lagged endogenous variables, $q_t'=(d_{t-1}',\dots,d_{t-p}')$ is a $qp$-dimensional vector of lagged deterministic (exogenous) variables, and $\Gamma_d=\begin{pmatrix} A_1',\dots,A_p'\end{pmatrix}$. The prior is


$$
\pi \begin{pmatrix} \Gamma_d, \Lambda, \Psi\end{pmatrix} = \pi \begin{pmatrix} \Gamma_d \end{pmatrix} \pi \begin{pmatrix} \Lambda \end{pmatrix} \pi \begin{pmatrix} \Psi \end{pmatrix}
$$


with $\pi \begin{pmatrix} \Gamma_d \end{pmatrix}$ and $\pi \begin{pmatrix} \Lambda \end{pmatrix}$ normal,


$$
\begin{aligned}
\gamma_d &\sim N(\underline{\gamma}_d, \underline{\Sigma}_d)\\
\lambda &= \textrm{vec} (\Lambda) \sim N(\underline{\lambda}, \underline{\Sigma}_{\lambda})
\end{aligned}
$$


and a Jeffreys' prior for $\Psi$. Alternatively a proper inverse Wishart, $\Psi \sim iW(\underline{S}, \underline{v})$, for $\Psi$ can be used.

Here $\pi \begin{pmatrix} \Gamma_d \end{pmatrix}$ is based on the Minnesota prior with overall tightness $\pi_1$, cross-equation tightness $\pi_2$ and lag decay rate $\pi_3$.


```{r, fig.width=10, fig.height=8,eval=FALSE}
rm(list = ls())
library(Karlsson2013inR)

data("villani2009")
yt <- villani2009
yt <- ts(yt[1:102, ], start = start(yt), frequency = frequency(yt))
plot.ts(yt)

bvar_obj <- bvar(data = yt)

bp <- which(time(yt) == 1992.75)
dum_var <- c(rep(1,bp), rep(0,nrow(yt)-bp))

bvar_obj <- setup(bvar_obj,
                  p=4,
                  deterministic = "constant_and_dummy",
                  dummy = dum_var)
pi_1 <- 0.2
pi_2 <- 0.5
pi_3 <- 1.0

#fol_pm = first own lag prior means
fol_pm=c(0,   #delta y_f
         0,   #pi_f
         0.9, #i_f
         0,   #delta y
         0,   #pi
         0.9, #i
         0.9  #q
         )

#lambda_1 = Lambda col 1
#lambda_2 = Lambda col 2
#lambda = vec(Lambda)

lambda_pr_mean <- 
  c(
  ppi( 2.00,  3.00,  annualized_growthrate=TRUE)$mean,   #lambda_1: delta y_f
  ppi( 1.50,  2.50,  annualized_growthrate=TRUE)$mean,   #lambda_1: pi_f
  ppi( 4.50,  5.50,  annualized_growthrate=FALSE)$mean,  #lambda_1: i_f
  ppi( 2.00,  2.50,  annualized_growthrate=TRUE)$mean,   #lambda_1: delta y
  ppi( 1.70,  2.30,  annualized_growthrate=TRUE)$mean,   #lambda_1: pi
  ppi( 4.00,  4.50,  annualized_growthrate=FALSE)$mean,  #lambda_1: i
  ppi( 3.85,  4.00,  annualized_growthrate=FALSE)$mean,  #lambda_1: q
  ppi(-1.00,  1.00,  annualized_growthrate=TRUE)$mean,   #lambda_2: delta y_f
  ppi( 1.50,  2.50,  annualized_growthrate=TRUE)$mean,   #lambda_2: pi_f
  ppi( 1.50,  2.50,  annualized_growthrate=FALSE)$mean,  #lambda_2: i_f
  ppi(-1.00,  1.00,  annualized_growthrate=TRUE)$mean,   #lambda_2: delta y
  ppi( 4.30,  5.70,  annualized_growthrate=TRUE)$mean,   #lambda_2: pi
  ppi( 3.00,  5.50,  annualized_growthrate=FALSE)$mean,  #lambda_2: i
  ppi(-0.50,  0.50,  annualized_growthrate=FALSE)$mean   #lambda_2: q
  )

lambda_pr_covmat <- 
  diag(
  c(
  ppi( 2.00,  3.00,  annualized_growthrate=TRUE)$var,    #lambda_1: delta y_f
  ppi( 1.50,  2.50,  annualized_growthrate=TRUE)$var,    #lambda_1: pi_f
  ppi( 4.50,  5.50,  annualized_growthrate=FALSE)$var,   #lambda_1: i_f
  ppi( 2.00,  2.50,  annualized_growthrate=TRUE)$var,    #lambda_1: delta y
  ppi( 1.70,  2.30,  annualized_growthrate=TRUE)$var,    #lambda_1: pi
  ppi( 4.00,  4.50,  annualized_growthrate=FALSE)$var,   #lambda_1: i
  ppi( 3.85,  4.00,  annualized_growthrate=FALSE)$var,   #lambda_1: q
  ppi(-1.00,  1.00,  annualized_growthrate=TRUE)$var,    #lambda_2: delta y_f
  ppi( 1.50,  2.50,  annualized_growthrate=TRUE)$var,    #lambda_2: pi_f
  ppi( 1.50,  2.50,  annualized_growthrate=FALSE)$var,   #lambda_2: i_f
  ppi(-1.00,  1.00,  annualized_growthrate=TRUE)$var,    #lambda_2: delta y
  ppi( 4.30,  5.70,  annualized_growthrate=TRUE)$var,    #lambda_2: pi
  ppi( 3.00,  5.50,  annualized_growthrate=FALSE)$var,   #lambda_2: i
  ppi(-0.50,  0.50,  annualized_growthrate=FALSE)$var    #lambda_2: q
  )
  )

bvar_obj <- priors(bvar_obj,
                   pi_1,
                   pi_2,
                   pi_3,
                   fol_pm,
                   lambda_pr_mean,
                   lambda_pr_covmat,
                   Jeffrey=TRUE)

p <- bvar_obj$setup$p
m <- bvar_obj$setup$m
mf <- 3 #first 3 variables are foreign in yt
restriction_matrix <- matrix(1, m*p, m)

for(i in 1:p){
  rows <- ((i-1)*m + mf + 1) : (i*m)
  cols <- 1:mf
  restriction_matrix[rows, cols] <- 0
}
bvar_obj <- restrict_Gamma_d(bvar_obj, restriction_matrix)

bvar_obj$predict$H <- 12
bvar_obj$predict$x_pred <- cbind(rep(1, 12), 0)

bvar_obj <- fit(bvar_obj,
                iter = 4000,
                warmup = 1000)

fcst <- forecast(bvar_obj,
                 ci = 0.95,
                 fcst_type = "mean",
                 growth_rate_idx = c(4,5),
                 plot_idx = c(4,5,6))
```